FINE SCALE REGISTRATION OF SIDEWALKS AND OTHER RIBBON-LIKE FEATURES FROM AERIAL IMAGERY

### 2018-08-20 Chapters/4.Expect_results/Results.tex
Updates from ShareLaTeX

We developed a six-step adopt approach for sidewalk segmentation. 
With the input \ac{VHR} imagery, we locate the sidewalk geometric information and it's initial trajectory, then we generate ribbon image on each sidewalk which has been locate in the map data. 
After that, we adopt density estimation on each pixel in the ribbon image. 
We apply the dynamic programming algorithm with fully developed penalty function to get precious boundaries on sidewalks, then just simply converted the ribbon image back to it's original shape.

\section{Validate Initial Trajectory}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/arizon_center.png}
    \caption[Demonstration on Arizona Center]{Detail demonstration on part of Arizona center, all sidewalks' \ac{GIS} information shows in Orange.}
    \label{fig:arizon_center}
\end{figure}

Before applying deep learning network to determine sidewalk's location, it's necessary to hand marking all sidewalk from imagery or extract the geometric information of each sidewalk from resources like \ac{OSM}. 
It's challenging since the geometric accuracy was not precise, which could cause performance issue since we rely on this information to segment the sidewalk part from maps for the next step. 
To ensure data accuracy, as shows in figure \ref{fig:arizon_center}, we spent days marking all sidewalks and their precise boundaries in Arizona Central for initial experiments. 
We can locate each sidewalk by using the geometric information and saved them into separate files. 
We also calculate the average sidewalk width from boundaries to decide reasonable window size in the following step.

\section{Ribbon Image Generation}

To apply the hypothesis, we develop a way to convert the given sidewalk into a ribbon image. 
As shown in figure \ref{fig:StraightenProcess}, we use the previous outcome (A) from the last step as input and connected all coordinate points data as the mid-line for the sample sidewalk. 
Then we apply a smooth function that we created to smooth the mid-line, so the shape of mid-line fits better to the actual sidewalk shape(B). 
We generate two parallel line as offsets on each side of the mid-line with a distance of twice the average width in between. 
Ideally even the mid-line is not perfectly laying on the middle, we would still fit the whole sidewalk into the ribbon image. 
For each point on the mid-line, we connect the corresponding points on both offsets to get the movement of the sidewalk and recorded them(C, D), we generate the straightened ribbon image by using the color data from each pixel in the points set that we record earlier(E). 
The image data in between each offset, with constant image width could be retrieved after that (F).
Then we reshape the image data to generate the ribbon image(G). 
Figure \ref{fig:Sample_Sidewalk_2} shows the complete ribbon-image output result on sample sidewalk. We separate it into parts to show better detail.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/straghten.pdf}
    \caption[Ribbon Image Generation]{Ribbon image generating process: from sub-figures A to G, we showed the step by step process to generate ribbon-image from a partial sidewalk. From raw input (A), we smoothed mid line and generated parallel off-site (B). Generated pixels (C) and connected each pixel to its correspond color data from the raw input (D). After gathered pixel data (E), we removed mid-line (F) and generated the ribbon image(G).}
    \label{fig:StraightenProcess}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{Figures/sample2_demo.png}
    \caption[Sample Sidewalk 2]{A ribbon image straighten output on sample sidewalk, row 1 shows the original input, row 2 to 5 show the ribbon image output separate into parts.}
    \label{fig:Sample_Sidewalk_2}
\end{figure}

\section{Density Estimation on Pixels}

Before move on to dynamic programming algorithm, we need to give each pixels a initial score for classification, to classify each pixel is sidewalk or not. 
To achieve that, as shows in figure \ref{fig:GMM_Sample_2}, we applied a density estimation function to find the probability for each pixel in the ribbon image. 
For foreground density, we used grey-scale to determine the probability likely-hood for the sidewalk-feature, where white is more likely to be sidewalk and black is less likely. 
The background density is the opposite, which made black is more likely to be sidewalk and white is not. 
We use the log likely-hood function to calculate the foreground and background density and estimate density on pixels.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/GMM_SAMPLE2.png}
    \caption[Density Estimation on Sample Sidewalk]{A sample figure with density estimation applied on figure \ref{fig:Sample_Sidewalk_2}. Where white indicates pixels are more likely to be a sidewalk feature and black indicates pixels are more likely to be a non-sidewalk feature. We separate it into 4 parts from column 1 to 4.}
    \label{fig:GMM_Sample_2}
\end{figure}

\section{The Dynamic Programming Algorithm}

We use our dynamic programming algorithm on the ribbon image to find the max score for each pixel that had better probability to be sidewalk feature that we mentioned in Chapter 4. 
More importantly, we introduce a penalty control function that makes each horizontal movement in the edges or center cost more than vertical movement. Basically, for each line of pixels, if they think the edges or center should shift more horizontally than one pixel compares to the last line, we score them with more penalty.

Figure \ref{fig:penalty} shows our algorithm managed to separate the driveway and sidewalk under the same feature as penalty increase. So we could apply the dynamic programming solution to locate the precise boundaries of given ribbon image. The penalty control improved our performance significantly. Because of this, our approach could predict the sidewalk under the tree shadow and other shadows or obstacles, which bring us more accurate result compare to other methods.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/penalty.png}
    \caption[Penalty Process]{Comparison with penalty change, we tested a variety of variables to decide which variable had best performance for width control. With the penalty increased at a reasonable amount, our algorithm decided not to recognize a driveway as sidewalk walk, in order to keep the sidewalk width consistent, even though they are the same texture.}
    \label{fig:penalty}
\end{figure}

As shown in figure \ref{fig:sample_result2}, after we generate the result on ribbon image, just need to simply covert the ribbon image back to its original shape by reverse the method that generates the ribbon image.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.95\textwidth]{Figures/sample2_result.png}
    \caption[Desire Output on Sample Sidewalk]{Detail demonstration on desire result after applying our dynamic programming approach. Where row 1 shows the result after reshape, row 2 to 5 shows the boundaries we calculated on ribbon image from \ref{fig:Sample_Sidewalk_2}.}
    \label{fig:sample_result2}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.85\textwidth]{Figures/prf.png}
    \caption[Parameters Evaluation]{Detail comparison on $c_{bend}$, $c_{shrink}$ and $c_{expand}$. Where red line represent Recall, green line represent Precision and blue line represent F1. We generate these based on equations from Recall (6.2), Precision (6.1), F1 (6.3).}
    \label{fig:change_on_recall}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{Figures/4_comparison.pdf}
    \caption[Methods Comparison on Sample Sidewalk 1]{A sample result compared with three different methods and ours. In the order of grabcut, slic, active Contours and ours from left to right. Grabcut is unable to separate the tree shadow from actual sidewalk feature, same with slic and active contours. Our results are more likely to predict the accurate edges under obstacles.}
    \label{fig:Sample_4_compare}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.78\textwidth]{Figures/sample_sidewalk_1.png}
    \caption[Methods Comparison on Sample Sidewalk 2]{A sample results compared with three different methods and ours. In the order of active Contours, grabcut, slic, ours from left to right. Which shows samples of separating drive way and gutters.}
    \label{fig:Sample_2_compare}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.86\textwidth]{Figures/2_comparison_needed.png}
    \caption[Methods Comparison on Sample Sidewalk 3]{A sample results compared with three different methods and ours. In the order of grabcut, slic, active Contours, ours from left to right. Which shows samples of separating drive way and gutters.}
    \label{fig:Sample_3_compare}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/Oxford_success_complex.png}
    \caption[Sample Sidewalk 3]{Demonstration on sample sidewalks with our approach. We use Oxford Area that include more complex sidewalks. It shows that given samples are able to separate trees, building shadows and camouflage with adjacent materials.}
    \label{fig:oxford1}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/ny1.png}
    \caption[Sample Sidewalk 4]{Demonstration on sample sidewalks with our approach. We use New York area that include more complex sidewalks. Our approach is able to separate tree shadows and obstacles. It also has a decent performance on predicating boundaries on partial sidewalks that are fully covered by tree leaves.}
    \label{fig:ny1}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/ny3.png}
    \caption[Sample Sidewalk 6]{Demonstration on sample sidewalks with our approach. It shows a continues site but we separate it due to it's length. Our approach did recognized the sidewalk part under the tree top and predict it's precise boundaries.}
    \label{fig:ny3}
\end{figure}

>>>>

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{Figures/prf.png}
    \caption[Parameters Evaluation]{Detail comparison on $c_{bend}$, $c_{shrink}$ and $c_{expand}$. Where red line represent Recall, green line represent Precision and blue line represent F1. We generate these based on equations from Recall (6.2), Precision (6.1), F1 (6.3).}
    \label{fig:change_on_recall}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{Figures/4_comparison.pdf}
    \caption[Methods Comparison on Sample Sidewalk 1]{A sample result compared with three different methods and ours. In the order of grabcut, slic, active Contours and ours from left to right. Grabcut is unable to separate the tree shadow from actual sidewalk feature, same with slic and active contours. Our results are more likely to predict the accurate edges under obstacles.}
    \label{fig:Sample_4_compare}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.78\textwidth]{Figures/sample_sidewalk_1.png}
    \caption[Methods Comparison on Sample Sidewalk 2]{A sample results compared with three different methods and ours. In the order of active Contours, grabcut, slic, ours from left to right. Which shows samples of separating drive way and gutters.}
    \label{fig:Sample_2_compare}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.86\textwidth]{Figures/2_comparison_needed.png}
    \caption[Methods Comparison on Sample Sidewalk 3]{A sample results compared with three different methods and ours. In the order of grabcut, slic, active Contours, ours from left to right. Which shows samples of separating drive way and gutters.}
    \label{fig:Sample_3_compare}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/Oxford_success_complex.png}
    \caption[Sample Sidewalk 3]{Demonstration on sample sidewalks with our approach. We use Oxford Area that include more complex sidewalks. It shows that given samples are able to separate trees, building shadows and camouflage with adjacent materials.}
    \label{fig:oxford1}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/ny1.png}
    \caption[Sample Sidewalk 4]{Demonstration on sample sidewalks with our approach. We use New York area that include more complex sidewalks. Our approach is able to separate tree shadows and obstacles. It also has a decent performance on predicating boundaries on partial sidewalks that are fully covered by tree leaves.}
    \label{fig:ny1}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/ny3.png}
    \caption[Sample Sidewalk 6]{Demonstration on sample sidewalks with our approach. It shows a continues site but we separate it due to it's length. Our approach did recognized the sidewalk part under the tree top and predict it's precise boundaries.}
    \label{fig:ny3}
\end{figure}

\section{Failure Cases}

We produce some failure cases to show what we need to improve for the future work.

\begin{enumerate}
    \item Miss matching data set: Figure \ref{fig:oxford_fail_1} shows the failure case when the input map data and the trajectory data not match. The map data is older than the trajectory data, so the boundaries we generated is not yet existed on the map. Since the road structures changes over time, it's possible to get the data set in different time stamp. 
    \item Irregular Intersections: Figure \ref{fig:oxford_fail_2} shows the failure case when process irregular shape intersections. Our approach is able to generate the boundaries when the sidewalk is able to reshape into ribbon-image. Such as in figure \ref{fig:oxford_fail_2}, it's not yet able to produce the sidewalk boundaries with irregular shape intersection.
    \item Fully blocked sidewalk: \figref{fig:ny_fail_1} shows sample of sidewalks that fully blocked by obstacles, tree crown in our case. Base on the initial trajectory we find from open source, there must be a sidewalk under the tree crown. We count this as 'failure' cases since it's impossible for us to determine the accuracy for our result. 
\end{enumerate}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.83\textwidth]{Figures/oxford_fail_1.png}
    \caption[Failure Case 1]{Demonstration on failure cases. Due to the fact that road structures may change overtime. It's possible to produce the output with the input map and the initial trajectory within different time stamp. Sample figure shows our approach produce 'incorrect' sidewalk boundaries because the map data is older than the trajectory so we are producing the sidewalk that were not built yet.}
    \label{fig:oxford_fail_1}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{Figures/oxford_fail_2.png}
    \caption[Failure Case 2]{Demonstration on failure cases. Our approach was developed to produce boundaries for single sidewalk, and it's not yet able to process intersection with irregular shape.}
    \label{fig:oxford_fail_2}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/ny_failcase.png}
    \caption[Failure Case 3]{Demonstration on failure cases. Our approach is not able to determine the sidewalk boundaries that's completely blocked by obstacles. The result is base on the average of sidewalks width but it's impossible for us to compare the prediction with accurate result.}
    \label{fig:ny_fail_1}
\end{figure}

\section{Future Work}

There's couple more ideas we would like to attempt.

\begin{enumerate}
    \item To simplify the process, we would like to generate the sidewalk with just latitudes and longitudes.

    we can use the latitudes and longitudes data to locate a part of a map via Overpass Turbo \cite{overpass_turbo}, it supports features such as sidewalks geometric location since it's connecting with the database from \ac{OSM} \cite{OpenStreetMap}. Also, with the input latitudes and longitudes data, we can extract the map with given range from Google map or other open map source. Then just simply pass in both data set with our approach to generate whole sidewalk information for the whole site.

    \item We'll give hypothesis to process the intersection area. We identify the point that is overlapped as intersection part (same node in multiple ways), we connect nodes on each way that are adjacent to the point. By extend each path a little bit more into the adjacent path to finish the intersection. Figure \ref{fig:intersection} shows the basic idea of our hypothesis, the sidewalk from the left merges into the left. It's connecting the left sidewalk into the right one and processed it as the same part.

\end{enumerate}

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{Figures/intersection.png}
    \caption[Intersection Demonstration]{A hypothesis demonstration on intersection. By extend each path a little bit more into the next adjacent path, we can connect them we a little 'elbow'.}
    \label{fig:intersection}
\end{figure}

### 2018-08-20 Chapters/5.Conclusion/Conclusion.tex
Updates from ShareLaTeX

\chapter{Conclusion}
In this thesis defense, a method to estimate the thickness and to refine the trajectory of a ribbon-like feature in \ac{VHR} aeral orthoimagery has been presented. 
We were able to obtain both high qualitative and quantitative results for segmenting walking paths in aerial imagery by refining \ac{OSM} features. The method is fast, and has only a few parameters with clear interpretations.
The proposed approach does not rely on extensive training on annotated data, but instead learns to fit the underlying imagery based on a trimap that is specified by parameters $\MinRadius{}, \MaxRadius{}, \MaxDistance{}$ which values that are easily determined based on the task, and are unlikely to require adjustment.

We evaluated this work by comparing it mainly to approaches for interactive image segmentation. 
Because most automatic approaches for extracting street-maps focus on extracting coarse trajectories (which we take as input) rather than on precise alignment and width estimation. 
We focused on modeling footways which are small, thin, easily occluded, and often modified due to construction.
We believe that this approach is a natural tool to improve segmentation by post-processing automatically extracted street networks, and also by enabling fast interactive annotation of aerial images by refining hand-entered paths such as those found in \ac{OSM}.

@@

\section{Failure Cases}

We produce some failure cases to show what we need to improve for the future work.

\begin{enumerate}
    \item Miss matching data set: Figure \ref{fig:oxford_fail_1} shows the failure case when the input map data and the trajectory data not match. The map data is older than the trajectory data, so the boundaries we generated is not yet existed on the map. Since the road structures changes over time, it's possible to get the data set in different time stamp. 
    \item Irregular Intersections: Figure \ref{fig:oxford_fail_2} shows the failure case when process irregular shape intersections. Our approach is able to generate the boundaries when the sidewalk is able to reshape into ribbon-image. Such as in figure \ref{fig:oxford_fail_2}, it's not yet able to produce the sidewalk boundaries with irregular shape intersection.
    \item Fully blocked sidewalk: \figref{fig:ny_fail_1} shows sample of sidewalks that fully blocked by obstacles, tree crown in our case. Base on the initial trajectory we find from open source, there must be a sidewalk under the tree crown. We count this as 'failure' cases since it's impossible for us to determine the accuracy for our result. 
\end{enumerate}

\begin{figure}
    \centering
    \includegraphics[width=0.83\textwidth]{Figures/oxford_fail_1.png}
    \caption[Failure Case 1]{Demonstration on failure cases. Due to the fact that road structures may change overtime. It's possible to produce the output with the input map and the initial trajectory within different time stamp. Sample figure shows our approach produce 'incorrect' sidewalk boundaries because the map data is older than the trajectory so we are producing the sidewalk that were not built yet.}
    \label{fig:oxford_fail_1}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{Figures/oxford_fail_2.png}
    \caption[Failure Case 2]{Demonstration on failure cases. Our approach was developed to produce boundaries for single sidewalk, and it's not yet able to process intersection with irregular shape.}
    \label{fig:oxford_fail_2}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/ny_failcase.png}
    \caption[Failure Case 3]{Demonstration on failure cases. Our approach is not able to determine the sidewalk boundaries that's completely blocked by obstacles. The result is base on the average of sidewalks width but it's impossible for us to compare the prediction with accurate result.}
    \label{fig:ny_fail_1}
\end{figure}

\section{Future Work}

There's couple more ideas we would like to attempt.

\begin{enumerate}
    \item To simplify the process, we would like to generate the sidewalk with just latitudes and longitudes.

    we can use the latitudes and longitudes data to locate a part of a map via Overpass Turbo \cite{overpass_turbo}, it supports features such as sidewalks geometric location since it's connecting with the database from \ac{OSM} \cite{OpenStreetMap}. Also, with the input latitudes and longitudes data, we can extract the map with given range from Google map or other open map source. Then just simply pass in both data set with our approach to generate whole sidewalk information for the whole site.

    \item We'll give hypothesis to process the intersection area. By identify the point that is overlapped as intersection part (same node in multiple ways), we connect nodes on each way that are adjacent to the point. By extend each path a little bit more into the adjacent path to finish the intersection.

\end{enumerate}

>>>>

\chapter{Conclusion}
In this thesis, a method to estimate the thickness and to refine the trajectory of a ribbon-like feature in \ac{VHR} aeral orthoimagery has been presented. 
We were able to obtain both high qualitative and quantitative results for segmenting walking paths in aerial imagery by refining \ac{OSM} features. 
The \ac{DPST} has only a few parameters with clear interpretations.
The proposed approach does not rely on extensive training on annotated data, but instead learns to fit the underlying imagery based on a trimap that is specified by parameters $\MinRadius{}, \MaxRadius{}, \MaxDistance{}$ which values that are easily determined based on the task, and are unlikely to require adjustment.

We evaluated this work by comparing it mainly to approaches for interactive image segmentation. 
Since most automatic approaches for extracting street-maps focus on extracting coarse trajectories (which we take as input) rather than on precise alignment and width estimation. 
We focused on modeling footways which are small, thin, easily occluded, and often modified due to construction.
We believe that this approach is a natural tool to improve segmentation by post-processing automatically extracted street networks, and also by enabling fast interactive annotation of aerial images by refining hand-entered paths such as those found in \ac{OSM}.

@@

### 2018-08-20 Thesis.tex
Updates from ShareLaTeX

    We present a novel approach for segmenting ribbon-like features, such as sidewalks. With the only knowledge of the approximate locations and thicknesses of thin features, this approach can register coarsely digitized footways to high-resolution aerial imagery with high accuracy. The key innovation is a dynamic-programming solution to search for the optimal set of thicknesses and offsets to an initial path. The method provides control over the rate of change of thickness and direction, and it does not rely on prior knowledge of the materials in the image. The approach will be helpful as an aid to digitization, and downstream applications such as registration of walking paths, canals, or roadways as well as for damage and change detection.

>>>>

    We present a novel approach for segmenting ribbon-like features, such as sidewalks. With only knowledge of the approximate locations and thicknesses of thin features, this approach can register coarsely digitized footways to high-resolution aerial imagery with high accuracy. The key innovation is a dynamic-programming solution to search for the optimal set of thicknesses and offsets to an initial path. The method provides control over the rate of change of thickness and direction, and does not rely on prior knowledge of the materials in the image. The approach will be helpful as an aid to digitization, and downstream applications such as registration of walking paths, canals, or roadways as well as for damage and change detection.
